Pulling oral and IV data from CvTdb

### 2025-08-12 data-raw/pulling_oral_iv_data_cvtdb.Rmd
Moved helper functions to CvTdb query report.

```{r}
library(DBI)
devtools::load_all() # So this loads the previous cvt object
library(dplyr)
library(tidyr)
library(readr)
```

@@

```{r quick_message_funtion}
quick_message <- function(text = character()) {
  stopifnot(length(text) > 0)

  cli::cli_par()
  cli::cli_text(paste("{.strong [i]}  ", text[1], sep = "\t"))
  text <- text[-1]

  for (this_text in seq_along(text)) {
    cli::cli_li(text[this_text])
  }
  cli::cli_end()
}

quick_stats <- function(.data, old_version = FALSE) {
  stopifnot(is.data.frame(.data))

  cli::cli_h2("Statistics on the current data")
  message(paste0("Total number of data points: ", nrow(.data)))
  message("Unique Species and data counts:")
  print(table(.data$species))

  message(
    "Number of unique studies: ",
    length(unique(.data$fk_study_id))
  )
  message(
    "Number of unique series: ",
    length(unique(.data$fk_series_id))
  )

  if (old_version) {
    message(
      "Number of unique chemicals: ",
      length(unique(.data$analyte_dtxsid)),
      "\n"
    )
    chem_by_medium <- .data[c("conc_medium_normalized", "analyte_dtxsid")] %>%
      filter(!is.na(analyte_dtxsid)) %>%
      distinct()
    chem_by_curset <- .data[c("curation_set_tag", "analyte_dtxsid")] %>%
      filter(!is.na(analyte_dtxsid)) %>%
      distinct()
  } else {
    message(
      "Number of unique chemicals: ",
      length(unique(.data$analyzed_chem_dtxsid)),
      "\n"
    )
    chem_by_medium <- .data[c("conc_medium_normalized", "analyzed_chem_dtxsid")] %>%
      filter(!is.na(analyzed_chem_dtxsid)) %>%
      distinct()
    chem_by_curset <- .data[c("curation_set_tag", "analyzed_chem_dtxsid")] %>%
      filter(!is.na(analyzed_chem_dtxsid)) %>%
      distinct()
  }

  message(
    "Unique Dose units: ",
    toString(unique(.data$dose_level_units_original)),
    "\n"
  )
  if (all(is.na(.data$conc_units_normalized))) {
    message(
      "Unique Concentration units: ",
      toString(unique(.data$conc_units_original)),
      "\n"
    )
  } else {
    message(
      "Unique Concentration units: ",
      toString(unique(.data$conc_units_normalized)),
      "\n"
    )
  }
  message("Data by unique route of administration:")
  print(table(.data$administration_route_normalized))
  message("Data by medium collected:")
  print(table(.data$conc_medium_normalized))
  message("Data by curation set:")
  print(table(.data$curation_set_tag))
  message("\n\nUnique chemicals per medium:")
  print(table(chem_by_medium$conc_medium_normalized))
  message("Unique chemicals per curation set:")
  print(table(chem_by_curset$curation_set_tag))
}

@@

unit_values_cols <- rlang::quos(
  conc_original, conc,
  conc_units_original, conc_units_normalized,
  conc_sd_original, conc_sd,
  loq, loq_units,
  dose_level_original, dose_level_normalized,
  dose_level_units_original, dose_volume_units,
  dose_volume, weight_kg
)

```{r diff_log_function}
# Now I want to write a quick function that will log relevant changes
# from previous 'cvt' objects
diff_log_cvt <- function(new_cvt,
                         key_id = "conc_time_id",
                         grp_ids = c(
                           "analyzed_chem_dtxsid",
                           "species"
                         ),
                         notes = character(0L)) {
  stopifnot(
    key_id %in% colnames(cvt),
    key_id %in% colnames(new_cvt),
    all(grp_ids %in% colnames(cvt)),
    all(grp_ids %in% colnames(new_cvt))
  )
  if (length(unique(new_cvt[[key_id]])) != NROW(new_cvt)) {
    stop("key_id is not not a column that uniquely identifies this data.")
  }

  # Prepare some vectors:
  grp_strings_new <- unique(do.call("paste", new_cvt[grp_ids]))
  grp_strings_old <- unique(do.call("paste", cvt[grp_ids]))

  msg_to_write <- paste0("\n\n======== ", round(Sys.time()))

  if (identical(new_cvt, cvt)) {
    msg_to_write <- paste(msg_to_write, "No changes.", sep = "\t")
    message(
      "There are no changes from previous `cvt` data.frame, ",
      "are you sure you want to save this?"
    )
  } else {
    msg_to_write <- paste(msg_to_write, "Changes detected.", sep = "\t")

    ### ---- Check key_id changes (usually conc_time_id)
    diff_row_add <- setdiff(new_cvt[[key_id]], cvt[[key_id]])
    diff_row_sub <- setdiff(cvt[[key_id]], new_cvt[[key_id]])
    diff_row_common <- intersect(cvt[[key_id]], new_cvt[[key_id]])
    # Nature of data changes:
    added_grps <- unique(
      do.call(
        "paste",
        new_cvt[which(new_cvt[[key_id]] %in% diff_row_add), grp_ids]
      )
    )
    removed_grps <- unique(
      do.call(
        "paste",
        cvt[which(cvt[[key_id]] %in% diff_row_sub), grp_ids]
      )
    )

    added_new_grps <- setdiff(added_grps, grp_strings_old)
    added_existing_grps <- intersect(added_grps, grp_strings_old)
    removed_new_grps <- setdiff(removed_grps, grp_strings_new)
    removed_existing_grps <- intersect(removed_grps, grp_strings_new)

    msg_to_write <- c(
      msg_to_write,
      paste("Rows added:", length(diff_row_add)),
      paste("Rows removed:", length(diff_row_sub))
    )

    limit_grp_output <- function(grp_vector) {
      if (length(grp_vector) > 0) {
        if (length(grp_vector) > 6) {
          grp_vector <- c(
            grp_vector[c(1, 2)],
            paste0(
              "... [",
              length(grp_vector) - 3,
              "] more!"
            ),
            grp_vector[c(length(grp_vector) - 1)]
          )
        }
        return(grp_vector)
      }
      msg_to_write <- c(
        msg_to_write,
        "New Groups added (not in previous data):",
        paste0("\t", limit_grp_output(added_new_grps)),
        "Groups with new observations (group present in previous data)",
        paste0("\t", limit_grp_output(added_existing_grps)),
        "Groups removed that no longer have data present:",
        paste0("\t", limit_grp_output(removed_new_grps)),
        "Groups removed that still have some data present:",
        paste0("\t", limit_grp_output(removed_existing_grps)),
        "-----------------------------------"
      )
    }

    ### -----Checking columns
    diff_col_add <- setdiff(names(new_cvt), names(cvt)) # Columns added in new_cvt
    diff_col_sub <- setdiff(names(cvt), names(new_cvt)) # Columns removed in new_cvt
    diff_col_common <- intersect(names(cvt), names(new_cvt))
    common_row_cvt <- cvt[cvt[[key_id]] %in% diff_row_common, ]

    common_row_cvt <- common_row_cvt[order(common_row_cvt[[key_id]]), ]

    common_row_new <- new_cvt[new_cvt[[key_id]] %in% diff_row_common, ]
    common_row_new <- common_row_new[order(common_row_new[[key_id]]), ]

    if (length(c(diff_col_add, diff_col_sub)) == 0) {
      msg_to_write <- c(
        msg_to_write,
        "No changes in column names."
      )
    } else if (length(diff_col_add) > 0 && length(diff_col_sub) > 0) {
      # Check renaming

      diff_columns <- function(x, y) {
        out_list <- character()
        for (ix in setdiff(names(x), key_id)) {
          for (iy in setdiff(names(y), key_id)) {
            tmp <- merge(x[c(key_id, ix)], y[c(key_id, iy)], by = key_id)
            if (identical(tmp[[2]], tmp[[3]])) {
              out_list <- c(out_list, paste0(ix, " -> ", iy))
            }
          }
        }
        return(out_list)
      }

      renamed_cols <- diff_columns(
        common_row_cvt[c(key_id, diff_col_sub)],
        common_row_new[c(key_id, diff_col_add)]
      )
      if (length(renamed_cols) > 0) {
        msg_to_write <- c(
          msg_to_write,
          "Renamed columns (non-exhaustive):",
          renamed_cols
        )
      } else {
        msg_to_write <- c(
          msg_to_write,
          "No renamed columns found."
        )
      }
    }
    if (length(diff_col_add) > 0) {
      msg_to_write <- c(
        msg_to_write,
        "--------------",
        "Columns added:",
        diff_col_add
      )
    }
    if (length(diff_col_sub) > 0) {
      msg_to_write <- c(
        msg_to_write,
        "______________",
        "Columns removed (from previous data):",
        diff_col_sub
      )
    }

    if ((NROW(
      setdiff(
        common_row_cvt[diff_col_common], common_row_new[diff_col_common]
      )
    ) > 0
    ) || (NROW(
      setdiff(
        common_row_new[diff_col_common], common_row_cvt[diff_col_common]
      )
    ) > 0
    )) {
      msg_to_write <- c(
        msg_to_write,
        "Data for same observation key_ids have changed."
      )

      # Which columns have different data?
      test_cols <- setdiff(diff_col_common, key_id)
      for (i in test_cols) {
        this_diff <- setdiff(
          common_row_new[c(key_id, i)],
          common_row_cvt[c(key_id, i)]
        )
        that_diff <- setdiff(
          common_row_cvt[c(key_id, i)],
          common_row_new[c(key_id, i)]
        )

        if (NROW(this_diff) > 0) {
          msg_to_write <- c(
            msg_to_write,
            paste0(c(i, "had", NROW(this_diff), "differences."), collapse = " ")
          )

          this_grp <- dplyr::distinct(common_row_new[
            common_row_new[[key_id]] %in% this_diff[[key_id]],
            grp_ids
          ])
          this_grp <- paste0(this_grp[1, 1:2], collapse = " ")

          # Print the differences (if below a certain number of rows)
          msg_to_write <- c(
            msg_to_write,
            paste0(
              "Example:",
              paste(this_grp, collapse = " "),
              collapse = " "
            ),
            paste0(c(i, "\t new:", this_diff[1, 2], "<--\t old:", that_diff[1, 2]),
              collapse = " "
            ), "\n"
          )
        }
      }
    }
    msg_to_write <- c(
      msg_to_write,
      notes,
      "================================\n"
    )
  }
  return(cat(msg_to_write, sep = "\n"))
}
```

@@

audit_query <- query_db(
  paste0(
    "SELECT DISTINCT * FROM cvt_audit ",
    "WHERE fk_table_name IN ('documents', 'series', 'studies')"
  ),
  "cvt"
)
study_query <- query_db(study_docs, "cvt")
series_query <- query_db(series_chems, "cvt")
dtxsid_query <- query_db("SELECT DISTINCT * FROM chemicals", "cvt") %>%
  mutate(
    across(ends_with("_id"), as.character),
    across(contains("_name"), \(x) {
      stringr::str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        stringr::str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        stringr::str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        stringr::str_replace_all("\\[3\\]h", "[3h]") |>
        stringr::str_replace_all("benzo\\[a\\]", "benzo(a)") |>
        # adds a space after square bracket separator
        stringr::str_replace_all("](?!\\s|-)", "] ")
    })
  )

@@

# Let's parse through with those "Some Identical" identifiers first
unequal_chem_ids |>
  filter(
    analyzed_dsstox_substance_id == dosed_dsstox_substance_id |
      analyzed_dsstox_casrn == dosed_dsstox_casrn |
      analyzed_preferred_name == dosed_preferred_name |
      analyzed_chemical_name_original == dosed_chemical_name_original |
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
  ) |>
  group_by(fk_analyzed_chemical_id, fk_dosed_chemical_id) |>
  summarize(
    same_dtxsid = if_else(
      analyzed_dsstox_substance_id == dosed_dsstox_substance_id,
      analyzed_dsstox_substance_id, NA_character_, missing = NA_character_
    ),
    same_casrn = if_else(
      analyzed_dsstox_casrn == dosed_dsstox_casrn,
      analyzed_dsstox_casrn, NA_character_, NA_character_
    ),
    same_preferred_name = if_else(
      analyzed_preferred_name == dosed_preferred_name,
      analyzed_preferred_name, NA_character_, NA_character_
    ),
    same_chem_name = if_else(
      analyzed_chemical_name_original == dosed_chemical_name_original,
      analyzed_chemical_name_original, NA_character_, NA_character_
    ),
    same_chem_name_secondary = if_else(
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original,
      analyzed_chemical_name_secondary_original, NA_character_, NA_character_
    )
  ) |>
  rowwise() |>
  mutate(
    likely_same = sum(!is.na(c_across(starts_with("same_"))))
  ) |>
  ungroup() |>
  arrange(same_dtxsid, same_casrn, same_preferred_name, same_chem_name,
          same_chem_name_secondary) |>
  View("Similar IDs")
# There are two ways the identifiers may be duplicated,
# shown here between rows and columns

@@

```{r harmonizing}
load("data/cvtdb_original.rda")
sql_query <- cvtdb_original
rm(cvtdb_original)

# Harmonize values for easier analysis
cvt_df <- sql_query %>%
  mutate(
    # Note that sometimes fk_test_chemical_id is NA
    across(all_of(c(
      "time_original",
      "time_hr",
      "conc_original",
      "conc_sd_original",
      "conc",
      "conc_sd",
      "dose_level_original",
      "dose_level_normalized",
      "loq"
    )), as.numeric),
    # conc, conc_sd_original, and conc_original
    # have NA/NE/ND values that are coerced to NAs
    year = as.character(year),
    species = tolower(species),
    sex = tolower(sex),
    across(contains("chem_name"), \(x) {
      stringr::str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        stringr::str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        stringr::str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        stringr::str_replace_all("\\[3\\]h", "[3h]") |>
        stringr::str_replace_all("\\[a\\]pyr", "(a)pyr") |>
        stringr::str_remove_all(" \\(aids\\)$") |>
        # adds a space after square bracket separator
        stringr::str_replace_all("](?!\\s|-)", "] ")
    }),
    dose_volume_units = ifelse(
      dose_volume_units %in% c("NR", "ND"),
      NA_character_,
      gsub(" / ", "/", gsub("ml", "mL", dose_volume_units))
    ),
    dose_volume = signif(as.numeric(dose_volume), digits = 7),
    dose_duration_units = trimws(dose_duration_units),
    dose_duration = as.numeric(stringr::str_extract(dose_duration,
      pattern = "^\\d+"
    )),
    dose_frequency_normalized = "Once", # Showa is not annotated
    radiolabeled = ifelse(is.na(radiolabeled), FALSE, as.logical(radiolabeled)),
    # for n_subject ranges, take integer value of median
    # otherwise pull number or NA
    n_subjects_normalized = case_when(
      n_subjects_in_series %in% "NR" ~ NA_real_,
      grepl("-", n_subjects_in_series) ~ floor(mean(
        as.numeric(unlist(strsplit(n_subjects_in_series, split = "-"))),
        na.rm = TRUE
      )),
      .default = as.numeric(
        stringr::str_extract(n_subjects_in_series, pattern = "\\d+")
      )
    )
  )
```

@@

There is also some radiolabeling experiments that aren't described as such.
We can use `stringr::str_detect(x,"(13|14)c")` to recode `radiolabeled = TRUE`
for these.

Then we should address the fact that the `invivoPKfit_2.0.0` version of the `cvt`
object has 228 chemicals, while the recent query only has 145. This is a
difference of 83.

@@

cli::cli_inform(
  "{length(unique(cvt_df$analyzed_chem_dtxsid))} unique analyzed_dtxsids"
)

cli::cli_inform(
  "{length(unique(cvt_df$dosed_chem_dtxsid))} unique dosed dtxsids"
)

# Fix Radiolabeled ones!
cvt_df %>%
  distinct(analyzed_chem_name_original, dosed_chem_name_original, radiolabeled) %>%
  filter(!radiolabeled) %>%
  filter(stringr::str_detect(dosed_chem_name_original, "((13|14)c|3h)"))
cvt_df %>%
  distinct(analyzed_chem_name_original, dosed_chem_name_original, radiolabeled) %>%
  filter(!radiolabeled) %>%
  filter(stringr::str_detect(analyzed_chem_name_original, "((13|14)c|3h)"))

cvt_df <- cvt_df %>%
  mutate(
    radiolabeled = ifelse(
      stringr::str_detect(dosed_chem_name_original, "((13|14)c|3h)") |
        stringr::str_detect(analyzed_chem_name_original, "((13|14)c|3h)"),
      TRUE,
      radiolabeled
    )
  )
# There is a distinct challenge when handling radiolabelled data but these will be
# filtered out later for our purposes.
# Pre-emptive filtering to cut down workload
cvt_df <- filter(cvt_df, !(radiolabeled %in% TRUE))

cvt_df$analyzed_chem_name_original[
  grepl("\\[\\w\\]", cvt_df$analyzed_chem_name_original)
] |> unique()

cvt_df$dosed_chem_name_original[
  grepl("\\[\\w\\]", cvt_df$dosed_chem_name_original)
] |> unique()

@@

cvt_df <- cvt_df |>
  mutate(
    # fix misspellings
    analyzed_chem_name_original = str_replace_lst(
      analyzed_chem_name_original,
      c(
        "paracetmol" = "paracetamol",
        "deacetyl diltiazem" = "desacetyl diltaiazem",
        "deacetyl n-monodesmethyl diltiazem" = "desacetyl n-monodesmethyl diltiazem",
        "2,2',4,4',5,5'-hexachlorobiphenyl" = "2,4,5,2',4',5'-hexachlorobiphenyl",
        "valproate \\(vpa\\)" = "valproic acid",
        "clotiapine" = "clothiapine"
      )
    ),
    dosed_chem_name_original = str_replace_lst(
      dosed_chem_name_original,
      c(
        "paracetmol" = "paracetamol",
        "deacetyl diltiazem" = "desacetyl diltaiazem",
        "deacetyl n-monodesmethyl diltiazem" = "desacetyl n-monodesmethyl diltiazem",
        "2,2',4,4',5,5'-hexachlorobiphenyl" = "2,4,5,2',4',5'-hexachlorobiphenyl",
        "clotiapine" = "clothiapine"
      )
    ),
    # Missing DTXSIDs I found
    dosed_chem_dtxsid = case_match_lst(
      dosed_chem_name_original,
      c(
        "4,4'-sulfonylbisphenol-d8" = "DTXSID3022409",
        "(+-)-metoprolol tartrate" = "DTXSID9037248",
        "antipyrene" = "DTXSID6021117",
        "3'-azido-3'-deoxythymidine" = "DTXSID8020127",
        "2,4,5,2',4',5'-hexachlorobiphenyl" = "DTXSID2032180"
      ),
      dosed_chem_dtxsid
    ),
    analyzed_chem_dtxsid = case_match_lst(
      analyzed_chem_name_original,
      c(
        "fluorotelomer acid 7+3" = "DTXSID20874028",
        "valproic acid" = "DTXSID6023733",
        "s-(-)-metoprolol" = "DTXSID60230861",
        "r-(+)-metoprolol" = "DTXSID20230862",
        "paracetamol" = "DTXSID2020006"
      ),
      analyzed_chem_dtxsid
    ),
    # check name matches
    analyzed_chem_dtxsid = ifelse(
      is.na(analyzed_chem_dtxsid) &
        analyzed_chem_name_original == dosed_chem_name_original,
      dosed_chem_dtxsid,
      analyzed_chem_dtxsid
    )
  )

quick_stats(cvt_df)
# quick_stats(cvt_2.0.0, old_version = FALSE)
# Note unique chemical counts

cvt_df |>
  distinct(!!!chem_id_cols) |> 
  filter(is.na(analyzed_chem_dtxsid)) |>
  View("ID-less")
# Anything without an ID for analyzed_chemical is likely a metabolite/control of sorts
# So... I filter it out.
cvt_df <- cvt_df |> 
  filter(!is.na(analyzed_chem_dtxsid))

# Now we need to pre-emptively address the chemicals that have different
# IDs but are actually the same chemical or represent the same chemical during the experiment
# This includes primarily: SALTS and ENANTIOMERS
cvt_df %>%
  filter(
    fk_analyzed_chemical_id != fk_dosed_chemical_id,
    analyzed_chem_dtxsid != dosed_chem_dtxsid
  ) %>%
  distinct(
    analyzed_chem_name_original, dosed_chem_name_original,
    fk_analyzed_chemical_id, fk_dosed_chemical_id,
    analyzed_chem_dtxsid, dosed_chem_dtxsid
  ) %>%
  arrange(analyzed_chem_name_original) %>%
  # head()
View()

>>>>

```{r}
library(DBI)
devtools::load_all() # So this loads the previous cvt object
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(cli)
```

@@

```{r quick_message_funtion}
quick_message <- function(text = character(0L)) {
  stopifnot(length(text) > 0)

  this_par <- cli_par()
  cli_text(text[1])
  text <- text[-1]

  ul <- cli_ul()
  for (this_text in text) {
    cli_li(this_text)
  }
  cli_end(ul)
  cli_end(this_par)
}

cli_print <- function(x, file = "", append = FALSE) {
  cat(x, file = file, sep = "\n", append = append)
}

# Wrapper for qcount())
qcount <- function(x) length(unique(x))

quick_stats <- function(.data) {
  stopifnot(is.data.frame(.data))

  .data <- subset(.data, !is.na(analyzed_chem_dtxsid))
  cli_text(c(
    "!" = "Only data with analyzed chemical dtxsid will be considered."
  ))
  cli_text("---- Statistics on the current data ----")
  cli_text("Total number of data points: {nrow(.data)}")
  cli_text("Unique Species and data counts:")
  cli_dl(table(.data$species))

  this_par <- cli_par()
  cli_text("Number of unique studies: {qcount(.data$fk_study_id)}")
  cli_text("Number of unique series: {qcount(.data$fk_series_id)}")
  cli_text("Number of unique chemicals: {qcount(.data$analyzed_chem_dtxsid)}")
  cli_end(this_par)

  that_par <- cli_par()
  cli_text("Original Dose Units: {qcount(.data$dose_level_units_original)}")
  cli_text("Normalized Dose Units: {qcount(.data$dose_level_units_normalized)}")
  cli_text("Original Conc Units: {qcount(.data$conc_units_original)}")
  cli_text("Normalized Conc Units: {qcount(.data$conc_units_normalized)}")
  cli_end(that_par)

  cli_par()
  cli_text("Data by unique route of administration:")
  cli_dl(table(.data$administration_route_normalized))
  cli_end()
  cli_par()
  cli_text("Data by unique medium collected:")
  cli_dl(table(.data$conc_medium_normalized))
  cli_end()
  cli_par()
  cli_text("Data by unique curation set:")
  cli_dl(table(.data$curation_set_tag))
  cli_end()
}

#' Wrapper for [dplyr::case_match()] that accepts named lists as arguments
#' @param .x A vector to match against. In a data context, a column.
#' @param .lst A named list in the form `old_value = new_value`.
#' @param .default Value to be used when there are no values. Suggest setting it to `.x`
#' @return A vector of replaced values
case_match_lst <- function(.x, .lst, .default = NULL) {
  stopifnot(rlang::is_named(.lst))

  dplyr::case_match(
    .x,
    !!!lapply(paste0("\"", names(.lst), "\"~\"", .lst, "\""), stats::as.formula),
    .default = .default
  )
}

#' Wrapper for [str_replace_all()] which replaces whole words.
#' @param .x A vector to match against.
#' @param .lst A named vector in the form `old_value = new_value`.
#' @return A character vector with values substituted in.
str_replace_lst <- function(.x, .lst) {
  stopifnot(rlang::is_named(.lst))

  str_replace_all(.x, setNames(.lst, paste0("^", names(.lst), "$")))
}

@@

unit_values_cols <- rlang::quos(
  conc_original, conc,
  conc_units_original, conc_units_normalized,
  conc_sd_original, conc_sd,
  loq, loq_units,
  dose_level_original, dose_level_normalized,
  dose_level_units_original, dose_volume_units,
  dose_volume, weight_kg
)
```

@@

audit_query <- query_db(
  paste0(
    "SELECT DISTINCT * FROM cvt_audit ",
    "WHERE fk_table_name IN ('documents', 'series', 'studies')"
  ),
  "cvt"
)
study_query <- query_dib(study_docs, "cvt")
series_query <- query_db(series_chems, "cvt")
dtxsid_query <- query_db("SELECT DISTINCT * FROM chemicals", "cvt") %>%
  mutate(
    across(ends_with("_id"), as.character),
    across(contains("_name"), \(x) {
      str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        str_replace_all("\\[3\\]h", "[3h]") |>
        str_replace_all("benzo\\[a\\]", "benzo(a)") |>
        # adds a space after square bracket separator
        str_replace_all("](?!\\s|-)", "] ")
    })
  )

@@

# Let's parse through with those "Some Identical" identifiers first
unequal_chem_ids |>
  filter(
    analyzed_dsstox_substance_id == dosed_dsstox_substance_id |
      analyzed_dsstox_casrn == dosed_dsstox_casrn |
      analyzed_preferred_name == dosed_preferred_name |
      analyzed_chemical_name_original == dosed_chemical_name_original |
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
  ) |>
  group_by(fk_analyzed_chemical_id, fk_dosed_chemical_id) |>
  summarize(
    same_dtxsid = if_else(
      analyzed_dsstox_substance_id == dosed_dsstox_substance_id,
      analyzed_dsstox_substance_id, NA_character_, missing = NA_character_
    ),
    same_casrn = if_else(
      analyzed_dsstox_casrn == dosed_dsstox_casrn,
      analyzed_dsstox_casrn, NA_character_, NA_character_
    ),
    same_preferred_name = if_else(
      analyzed_preferred_name == dosed_preferred_name,
      analyzed_preferred_name, NA_character_, NA_character_
    ),
    same_chem_name = if_else(
      analyzed_chemical_name_original == dosed_chemical_name_original,
      analyzed_chemical_name_original, NA_character_, NA_character_
    ),
    same_chem_name_secondary = if_else(
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original,
      analyzed_chemical_name_secondary_original, NA_character_, NA_character_
    )
  ) |>
  rowwise() |>
  mutate(
    likely_same = sum(!is.na(c_across(starts_with("same_"))))
  ) |>
  ungroup() |>
  arrange(
    same_dtxsid, same_casrn, same_preferred_name, same_chem_name,
    same_chem_name_secondary
  ) |>
  View("Similar IDs")
# There are two ways the identifiers may be duplicated,
# shown here between rows and columns

@@

```{r harmonizing}
log_file <- "data-raw/creating_cvt_object.txt"
cli_print(
  "Creating the `cvt` data object for invivoPKfit, after SQL query.",
  file = log_file,
  append = FALSE # Overwrite file
)

load("data/cvtdb_original.rda")
sql_query <- cvtdb_original
rm(cvtdb_original)

columns_to_numeric <- c(
  "time_original",
  "time_hr",
  "conc_original",
  "conc_sd_original",
  "conc",
  "conc_sd",
  "dose_level_original",
  "dose_level_normalized",
  "loq"
)

# Harmonize values for easier analysis
cvt_df <- sql_query %>%
  mutate(
    # Note that sometimes fk_test_chemical_id is NA
    across(all_of(columns_to_numeric), as.numeric),
    # conc, conc_sd_original, and conc_original
    # have NA/NE/ND values that are coerced to NAs
    year = as.character(year),
    species = tolower(species),
    sex = tolower(sex),
    across(contains("chem_name"), \(x) {
      str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        str_replace_all("\\[3\\]h", "[3h]") |>
        str_replace_all("\\[a\\]pyr", "(a)pyr") |>
        str_remove_all(" \\(aids\\)$") |>
        # adds a space after square bracket separator
        str_replace_all("](?!\\s|-)", "] ")
    }),
    dose_volume_units = ifelse(
      dose_volume_units %in% c("NR", "ND"),
      NA_character_,
      gsub(" / ", "/", gsub("ml", "mL", dose_volume_units))
    ),
    dose_volume = signif(as.numeric(dose_volume), digits = 7),
    dose_duration_units = trimws(dose_duration_units),
    dose_duration = as.numeric(str_extract(dose_duration,
      pattern = "^\\d+"
    )),
    dose_frequency_normalized = "Once", # Showa is not annotated
    radiolabeled = ifelse(is.na(radiolabeled), FALSE, as.logical(radiolabeled)),
    # for n_subject ranges, take integer value of median
    # otherwise pull number or NA
    n_subjects_normalized = case_when(
      n_subjects_in_series %in% "NR" ~ NA_real_,
      grepl("-", n_subjects_in_series) ~ floor(mean(
        as.numeric(unlist(strsplit(n_subjects_in_series, split = "-"))),
        na.rm = TRUE
      )),
      .default = as.numeric(
        str_extract(n_subjects_in_series, pattern = "\\d+")
      )
    )
  )

qm <- quick_message(c(
  "Column changes:",
  "Set {.var {columns_to_numeric}} to numeric type.",
  "Set {.var year} to character type.",
  "Set {.var {c('species', 'sex')}} to lowercase.",
  "Set {.var radiolabeled} to logical.",
  "Normalized some chemicals names. (See code for details)",
  "If `dose_volume_units` was 'NR' or 'ND' set this to {NA_character_}.",
  "Replaced ' / ', with '/' and 'ml' with 'mL' for `dose_volume_units`.",
  "Set `dose_duration` to first number and converted to numeric.",
  "Trimmed whitespace around `dose_duration_units`.",
  "Extracted integer average from `n_subjects_in_series` and set 'NR' to NA."
)) |> cli_fmt()

cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)
```

@@

There is also some radiolabeling experiments that aren't described as such.
We can use `str_detect(x,"(13|14)c")` to recode `radiolabeled = TRUE`
for these. Then I filter out radiolabeled data because currently _invivoPKfit_
does not handle these.

@@

cli_text("{qcount(cvt_df$analyzed_chem_dtxsid)} unique analyzed_dtxsids")

cli_text("{qcount(cvt_df$dosed_chem_dtxsid)} unique dosed dtxsids")

# Fix Radiolabeled ones!
cvt_df %>%
  distinct(
    fk_dosed_chemical_id,
    analyzed_chem_name_original,
    dosed_chem_name_original,
    radiolabeled
  ) %>%
  filter(
    !radiolabeled,
    str_detect(dosed_chem_name_original, "((13|14)c|3h)") |
      str_detect(analyzed_chem_name_original, "((13|14)c|3h)")
  )

cvt_df <- cvt_df %>%
  mutate(
    radiolabeled = ifelse(
      str_detect(dosed_chem_name_original, "((13|14)c|3h)") |
        str_detect(analyzed_chem_name_original, "((13|14)c|3h)"),
      TRUE,
      radiolabeled
    )
  )
# There is a distinct challenge when handling radiolabelled data but these will be
# filtered out later for our purposes.
# Pre-emptive filtering to cut down workload
cvt_df <- filter(cvt_df, !(radiolabeled %in% TRUE))

qm <- quick_message(c(
  "Found chemicals that were missidentified as not radiolabeled.",
  "Dosed Chemical IDs = (710, 1108)",
  "Filtered out radiolabeled data."
)) |> cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

@@

name_misspellings <- c(
  "paracetmol" = "paracetamol",
  "deacetyl diltiazem" = "desacetyl diltaiazem",
  "deacetyl n-monodesmethyl diltiazem" = "desacetyl n-monodesmethyl diltiazem",
  "2,2',4,4',5,5'-hexachlorobiphenyl" = "2,4,5,2',4',5'-hexachlorobiphenyl",
  "valproate \\(vpa\\)" = "valproic acid",
  "clotiapine" = "clothiapine"
)

found_dtxsids <- c(
  "4,4'-sulfonylbisphenol-d8" = "DTXSID3022409",
  "(+-)-metoprolol tartrate" = "DTXSID9037248",
  "antipyrene" = "DTXSID6021117",
  "3'-azido-3'-deoxythymidine" = "DTXSID8020127",
  "2,4,5,2',4',5'-hexachlorobiphenyl" = "DTXSID2032180",
  "fluorotelomer acid 7+3" = "DTXSID20874028",
  "valproic acid" = "DTXSID6023733",
  "s-(-)-metoprolol" = "DTXSID60230861",
  "r-(+)-metoprolol" = "DTXSID20230862",
  "paracetamol" = "DTXSID2020006"
)

cvt_df <- cvt_df |>
  mutate(
    # fix misspellings
    analyzed_chem_name_original = str_replace_lst(
      analyzed_chem_name_original,
      name_misspellings
    ),
    dosed_chem_name_original = str_replace_lst(
      dosed_chem_name_original,
      name_misspellings
    ),
    # Missing DTXSIDs I found
    dosed_chem_dtxsid = case_match_lst(
      dosed_chem_name_original,
      found_dtxsids,
      dosed_chem_dtxsid
    ),
    analyzed_chem_dtxsid = case_match_lst(
      analyzed_chem_name_original,
      found_dtxsids,
      analyzed_chem_dtxsid
    ),
    # check name matches
    analyzed_chem_dtxsid = ifelse(
      is.na(analyzed_chem_dtxsid) &
        analyzed_chem_name_original == dosed_chem_name_original,
      dosed_chem_dtxsid,
      analyzed_chem_dtxsid
    )
  )

qm <- cli_fmt({
  cli_par()
  cli_text("Fixed the following misspellings [old: new] :")
  cli_dl(name_misspellings)
  cli_par()
  cli_text("Added missing DTXSIDs:")
  cli_dl(found_dtxsids)
  cli_end()
})
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

quick_stats(cvt_df)
# Note unique chemical counts are

cvt_df |>
  distinct(!!!chem_id_cols) |>
  filter(is.na(analyzed_chem_dtxsid)) |>
  View("ID-less")

# Anything without an ID for analyzed_chemical is likely a metabolite/control of sorts
# Or justifiably undetermined...
# So... I filter it out.
cvt_df <- cvt_df |>
  filter(!is.na(analyzed_chem_dtxsid))

cli_print(
  cli_fmt({
    quick_message("Filtered out observations where `analyzed_chem_dtxsid` is still NA.")
  }),
  file = log_file,
  append = TRUE
)

Now we need to preemptively address the chemicals that have different
IDs but actually represent the same chemical during the experiment
This includes primarily: SALTS and stereoisomers

**Small note about stereoisomers:**  
Here I give an un-elaborate treatment to racemic mixtures,
both stereoisomers/enantiomers are assumed to be 100% of the dose.
This isn't physically possible but the assumption would be that for each 
analyzed stereoisomer, the dose is representing the maximum possible dose in those
cases. For the purposes of _invivoPKfit_ the complication lies in possible
overestimation of elimination or underestimation of absorption.
Since this is rare in the data, I note this caveat but keep the data.
